\chapter{Teacher-Led OurPlace Engagements}
\label{chap:Teachers}

\section{Overview}

As well as investigating how mobile learning platforms such as OurPlace could be used by community experts and stakeholders for sharing their knowledge and values, we were also interested in exploring if and how such tools could work within a formal education context. To this end, at the same time as working with the community heritage groups we were also working with teachers to investigate the use of ParkLearn as a seamless, place-based learning tool, garnering feedback as to how the application could be improved. This involved a longitudinal study with a local primary school, as well as several one-off studies with other schools in the area. The work covered by Section \ref{sec:LongitudinalSchool} was part of a body of work which was peer-reviewed and published at MobileHCI 2018 \citep{Richardson2018}, with the paper being co-authored by Doctors Pradthana Jarusriboonchai, Kyle Montague and Ahmed Kharrufa.

As these studies were running concurrently with the community engagements covered in Chapter \ref{chap:Community}, the application still existed as `ParkLearn', without the features and re-branding of `OurPlace'.

\section{A Longitudinal Study with a Primary School}
\label{sec:LongitudinalSchool}

I was put into contact with a teacher at a local primary school through a co-researcher, Dr Jarusriboonchai, who was working with them on a different research project. I worked with this teacher (Teacher 1, Year 4---aged 6-7) for a period of roughly one year, using the ParkLearn application multiple times on school trips, in the classroom and on the school's grounds. I also worked with another teacher (Teacher 2, Year 6---aged 10-11) in the same school, although this was limited due to examination pressures. The longer study period was chosen for two main reasons: to mitigate the influence of `novelty' in the children’s engagement with the technology (the hope was that once students had used the application multiple times, more authentic engagement could be observed---rather than simply excitement at the chance to use new technology) \citep{Sharples2013} and to see how the teachers' approaches to Activity creation would change as they gained experience with the application over time.

\subsection{Context}
The school is situated in one of the most economically deprived areas of England: the ward in which the school sits features the highest crime rate within the constituency, 26\% of its population are within the 10\% most deprived in the UK, and 15\% of the children within the ward live in poverty (down from 27\% in 2010). The area's life expectancy is 73 for people born male and 78 for people born female, significantly lower than the national average (around 88 years).

Despite this, the school itself is of a very high standard, having been awarded an `Outstanding' rating in their latest Ofsted (the UK’s Office for Standards in Education) review. While they don't have access to their own transportation (having to instead hire coaches), the school lies within a short driving distance of the park referred to in Section \ref{sec:TalkingStatue}. It also features its own grounds, including a tarmacked playground, sizable green space and a small wooded area. 

While Teacher 1 was not very confident with digital technology, Teacher 2 was seen as the school's `whizz-kid' teacher. As they taught a slightly older class of students, Teacher 2 frequently tasked them with doing online research during classroom activities. To this end, the school had within recent years established a partnership with Samsung, who had supplied them with a smart classroom display and 20 Android tablets, which were a shared resource amongst all the school’s classes (with older classes given priority).

We worked with the primary school (ages 4-11) over a period of four months to evaluate and further develop ParkLearn. We were working with two teachers: ‘Teacher 1’ who taught Year 6 (aged 10-11, we engaged N=16 children) and ‘Teacher 2’ who taught a Year 2 class (aged 6-7, N=29). As there were more children than devices in most instances, the tablets were usually shared one-between-two, with children taking turns to use the application at their own volition. The Year 6 class and teacher were less frequently available due to the study coinciding with exams. The teachers used the application at their own volition and took on a co-researcher role, creating the Activities independently and developing their own design ideas. While the Year 6 teacher frequently used tablet devices in the classroom and had a good level of technical experience, the Year 2 teacher had very little experience with smartphone applications. Prior to the application being used in class, we sat down with the teachers for an hour to give them a brief overview of the study, the application itself and how it could be used. To serve as examples for what the app could be used for, we had created two simple Activities beforehand: one which took the learner on a bug hunt, and a more creative Activity about movie making.

The teachers chose to use the application total of nine times between the two classes during the four-month study period: twice with the Year 6 class, and seven times with Year 2. The following observations were recorded in field notes.

\subsection{Introducing the Application}

The teachers both decided to hold introductory lessons with the app with their respective classes. Their first Activity was very exploratory, designed for use in the classrooms to see how easily the children could use the application (Table 2, row 2). The older group of children were already extremely comfortable with using them and had very few issues understanding the application’s design language. Some of the younger children were less able readers, and so struggled to understand even the simple instructions for each Action created by the teacher. To mitigate this, subsequent versions of the application featured the text-to-speech function, which read aloud the Action’s instruction at the push of a button.

\subsection{Year 6 Interactions}

The Year 6 group used the application on a trip to a site popular with school groups thanks to its historical, natural and scientific features. The site featured an indoor museum and a large outdoor property featuring woodlands and ornamental gardens. The teacher prepared the Activity the night before, using his prior knowledge of the location to design Actions which included: ‘Take Photos’ of the various wooden bridges present; ‘Photo Match’ Actions of a modern water pump and an iron bridge; a ‘Record Audio’ of the natural sounds of the forest; a ‘Record Video’ of an Archimedes screw rotating; a ‘Map Marking’ Action to plot where the powerhouse was; a ‘Location Hunt’ to navigate the children to a mystery location (an old waterwheel); and a final Action which challenged the children to compete and ‘Take Photos’ of the most beautiful flower they could find (Table 2, row 3). As the location was a significant distance away from the school, the teacher created the ‘Photo Match’ Actions using photographs downloaded from Google Images rather than taking them himself. Unfortunately, the ‘Map Marking’ Action wasn’t available to complete, due to its reliance on Internet connectivity to load Google Maps.

\subsection{Year 2 Interactions}

The Year 2 class used ParkLearn on three separate school trips, as well as during multiple activities on the school grounds. The first trip was to the park, with two different Activities: the first Activity focused on the historical monuments and memorials in the park, and asked the children to record each other explaining what they were dedicated to (Table 2, row 4); the second used ‘Photo Match’ Tasks to find and photograph specific flowers in the park, with a final ‘Take Photos’ Task asking them to choose their favourite (Table 2, row 5). The Year 2 teacher also used the application on a trip to another local park, where the ranger had invited her class to make suggestions as to how the park should be improved. The teacher made an Activity which asked her students to take photos of parts of the park and make audio recordings which would then be shared with the park ranger (Table 2, row 6). Unfortunately, a bug in this version of the app resulted in a loss of several children’s work, meaning that the children’s feedback was sent to the ranger as part of a classroom writing exercise instead. The Year 2 class also used the application to identify and talk about the trees they could find on the school grounds, through a mixture of Actions ranging from ‘Photo Matching’ trees to ‘Audio Recording’ how it felt to be around nature (Table 2, row 7). The teacher also used the application on a class trip to the zoo (Figure 1), with a simple Activity consisting of Actions which asked the children to take photos of each of the different animal types and record video and audio about things they found interesting (Table 2, row 8).

The final use of the app took place at the end of the school year. The teacher created an Activity which asked four children from her class to give advice to the younger Year 1 students about being in Year 2 (Table 2, row 9). The Activity asked the children to photograph an area of her classroom and record a video giving advice about the do’s and don’ts. The teacher chose the children either because they would benefit from the practice (due to a lack of confidence or, in the case of ‘Child 1’, a speech impediment) or because they were especially enthusiastic about using the application. Once the children’s responses were uploaded, the teacher played their videos for the class via the ParkLearn website.

\section{Other Adult-Led Engagements}

\subsection{Hawthorn Primary}

\subsection{Researcher-Led Engagements}

St Teresa's

\section{Observations of Teachers' Use of OurPlace}

We’ve structured our observation and interview data into the three themes that emerged from our thematic analysis:

\subsection{Supporting Seamless Learning Practices}

Children in both age groups were easily able to use the application to independently collect data, allowing them to make the most of being in the field: multimedia responses allowed immediate data collection and reflection, without struggling with poor writing skills and virtual keyboards. This was especially true with the younger children, many of whom weren’t strong writers. When asked in a later semi-structured interview, their teacher revealed that she purposefully chose Action Types which wouldn’t be technically challenging, allowing children to focus on the Tasks’ content: “It’s automatic. They can just speak. [...] When I designed the activity, I basically did the video, because I wanted them not to have to write.” The Year 2 teacher particularly favoured use of the camera, taking up over 80\% of her created Actions (Figure 3). The children’s interactions with the technology became more purposeful as the study progressed, unhindered by a lack of familiarity and the earlier versions’ bugs. This was shown in the trip to the zoo: the children were careful to correctly classify each animal into the correct Action, trying to take as good a photograph as possible and deliberately deleting and re-taking any shots that didn’t meet their increasingly high standards. One pair of children even decided to reshoot their video recording twice to ensure their delivery of information was perfect. Despite this perfectionism, each pair still uploaded over 14 photos on average in addition to their audio and video recordings (Table 2, row 8). The older children also responded well to their Activity, particularly enjoying competing to take the best flower photo, the ‘Location Hunt’ Action’s sound and animation and competing to take the most accurate ‘Photo Match’.

The final Year 2 Activity proved to be a very different use-case for the application: most of the learning process took place independent of the technology, as the children created their presentations and practiced with each other using whiteboards and markers. The application was used to prompt the children for talking points and to record their final output. The ability to prepare and redo a video presentation proved very effective for the children such as Child 1, who would have normally struggled due to a lack of confidence. Reflecting on the activity afterwards, the teacher stated that not only did the children enjoy recording the videos, but that they also took pride in the final results: “[Child 1 would present his work], but he doesn’t know what he’s going to say, he gets tongue-tied. The pride he’ll take in actually being able to give a coherent message and seeing himself back… They far more enjoyed what they were saying and what they were doing.” While two of the children didn’t want to play back their recordings for themselves, all four participating children were eager to show their videos to the rest of the class. The other children reacted with excitement at seeing their classmates on the screen, with Child 1 even receiving high-fives. 

The teachers saw great value in how simple the app made creating a structured learning activity and collecting the children’s responses to it. When interviewed towards the end of the study, Teacher 2 noted: “It’s powerful, really powerful. The way that packages it up at the end, and how immediate it is, is fantastic for me. I got it straight away, it wasn’t a difficult process to do.” Prior to the study, the school’s teachers had been manually transferring the children’s created photos and videos over USB on a weekly basis, uploading the children’s media to Earwig—an evidence portfolio suite used by the school. ParkLearn’s small output file sizes and upload system was far simpler and better suited to the teachers’ workflow: “That simplicity takes away a lot of hassle—if I was to take photographs on my [tablet], I’ve got to get the lead, plug it into my hard drive, transfer the photos across, choose where I want to save them... Whereas this packages everything together.” This one-step system took a fraction of the time compared to the old backup routine, meaning children’s creations would be discarded less often. Its simplicity even allowed the teachers to delegate uploading to the children. The Year 2 teacher valued that the submissions appeared on the website in the same format as they appeared on the learner’s device: “What I like about the app is you can pull together different ways of collecting and showing information. Simply by pressing that upload button, it puts it onto my screen to save and to use in that format. That’s the beauty of it.” Because the application supported both open-ended and structured learning activities and was non-intrusive in her workflow, she plans to use the application regularly with her next class of students. When asked which school trips would benefit from this data collection, she responded: “Every trip.”

\subsection{Engagement and Empowerment Through Ownership}

Both teachers created Activities which ranged between being highly prescribed (e.g. A ‘Photo Match’ with ‘Find a birch tree’) and open-ended (e.g. A ‘Take Photos’ with ‘Find what you think is the most beautiful flower’). The Year 2 class’s first Activity proved to be very prescriptive, with the teacher simply having the children repeat her words on video. While the children enjoyed recording each other with the tablets, they weren’t very engaged with the actual educational content (suggesting a high influence of the technology’s novelty factor). The Year 2 teacher noted in an interview that in these cases, the children were only really interested in viewing their own work: “When they come back after visits where we’ve all done the same, children’s enthusiasm is not really there for what others have done. The enthusiasm is, “Can I see what I’ve done?” The children took pride in the photos they had taken, showing off their creations to each other and the adults present. She started planning future Activities which would involve the children having their own topics in small groups: “They’re given a specific task and they take ownership of it, knowing that other groups are not doing that. […] When we come back to school and we feedback, there’s a great interest in what each other has produced because we’re informing everybody.” She argued that covering their own topics would lead to the children becoming experts on it amongst their friends, with the ownership of the task and knowledge empowering them through the ability to teach others: “You can kind of empower yourself through your knowledge and how you’re going to present it, and then go off and do it.”

\subsection{Supporting Civic Engagement and Inquiry}

The platform was used as a tool to facilitate civic participation during the class trip to the second park (Table 2, row 6), with students providing feedback and suggestions through the application. The Year 2 teacher noted that the school was struggling to engage parents in the children’s contributions: “We don’t have a lot of parental support, but, where we do, we’re looking to make cultural links.” After this trip to the park, she had tried to make the parents aware of their children’s work through the school newsletter. While their currently used technology, ‘Earwig’, was impractical for this (“It takes a very long time to upload a video within Earwig, so we’ve stopped doing it, really.”), she suggested that mobile technologies may be better suited to highlighting the children’s civic engagements: “Our parents might go along and just say, ‘There’s nothing there,’ because they don’t see the resource. However, there could be something on the app like, ‘We’re involved in it, so go along and see what your children have done.’” Other activities that she suggested could be highlighted included visits to local care homes: “Let’s say, Christmas you go to the care home. We can use ParkLearn to record what we did and use that within school, upload it to our website to share it with parents.” This highlighting could extend to cross-cultural learning engagements: “We have a link with a school in India. The app is a perfect way of interacting with them, showing each other.” The platform was also used to promote self-reflection on the learner’s relationship with the space. For example, one Activity asked children to record audio in response to the prompt: ‘Why are trees special? Listen to the sound of the leaves rustling, stand amongst them and look up – how does it make you feel? Share your thoughts with us.’ 

\section{Summary}